---
title: "p8105_hw6_ww2744"
author: "Wenjie Wu"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(broom)
library(ggplot2)
library(dplyr)
library(purrr)
library(modelr)
library(rsample)
```

# Problem 1

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
bootstrap_samples = bootstraps(weather_df, times = 5000)

calculate_metrics = function(split) {
  data = analysis(split)
  fit = lm(tmax ~ tmin, data = data)
  r_squared = glance(fit)$r.squared
  coefs = tidy(fit)
  beta0 = coefs$estimate[coefs$term == "(Intercept)"]
  beta1 = coefs$estimate[coefs$term == "tmin"]
  log_beta_product = log(beta0 * beta1)
  return(tibble(r_squared = r_squared, log_beta_product = log_beta_product))
}

bootstrap_results = bootstrap_samples |>
  mutate(metrics = map(splits, calculate_metrics)) |>
  unnest(metrics)

bootstrap_summary = bootstrap_results |>
  summarize(
    r2_lower = quantile(r_squared, 0.025),
    r2_upper = quantile(r_squared, 0.975),
    log_beta_lower = quantile(log_beta_product, 0.025),
    log_beta_upper = quantile(log_beta_product, 0.975)
  )

print(bootstrap_summary)

ggplot(bootstrap_results, aes(x = r_squared)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  labs(title = "Distribution of R-squared", x = "R-squared", y = "Frequency") +
  theme_minimal()

ggplot(bootstrap_results, aes(x = log_beta_product)) +
  geom_histogram(bins = 30, fill = "lightgreen", color = "black") +
  labs(title = "Distribution of log(beta0 * beta1)", x = "log(beta0 * beta1)", y = "Frequency") +
  theme_minimal()
```


# Problem 2

```{r}
homi_df = read_csv("data/homicide-data.csv")

homi_df = homi_df |>
  mutate(city_state = paste(city, state, sep = ", ")) |>
  mutate(solved = ifelse(grepl("Closed", disposition, ignore.case = TRUE), 1, 0))

omit_cities = c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")

homi_df = homi_df |>
  filter(!city_state %in% omit_cities) |>
  filter(victim_race %in% c("White", "Black")) |>
  mutate(victim_age = as.numeric(victim_age))
```

```{r}
baltimore_data = homi_df |>
  filter(city_state == "Baltimore, MD")

glm_model = glm(solved ~ victim_age + victim_sex + victim_race, 
                 data = baltimore_data, 
                 family = binomial)

summary = broom::tidy(glm_model, conf.int = TRUE, exponentiate = TRUE)
print(summary)

male_female = summary |>
  filter(term == "victim_sexMale") |>
  select(term, estimate, conf.low, conf.high)
print(male_female)

```

```{r}
nested_city = homi_df |>
  group_by(city_state) |>
  nest()

city_glm_results = nested_city |>
  mutate(glm_model = map(data, ~ glm(solved ~ victim_age + victim_sex + victim_race,
                                     data = ., family = binomial)),
         tidy_model = map(glm_model, ~ tidy(., conf.int = TRUE, exponentiate = TRUE))) |>
  select(city_state, tidy_model) |>
  unnest(tidy_model) |>
  filter(term == "victim_sexMale") |>
  select(city_state, estimate, conf.low, conf.high) |>
  arrange(estimate) |>
  mutate(city_state = factor(city_state, levels = city_state)) |>
  print()

```

```{r}
ggplot(city_glm_results, aes(x = fct_reorder(city_state, estimate), y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  coord_flip() +
  labs(title = "Adjusted Odds Ratios for Solving Homicides",
       x = "City",
       y = "Adjusted Odds Ratio") +
  theme_minimal(base_size = 10) + 
  theme(axis.text.y = element_text(size = 5))
```

- The variability in odds ratios across cities highlights differences in investigative outcomes，potentially influenced by local factors. The top 3 city has the lowest estimate odds ratio is `New York`, `Long Beach` and `Denver`. 

- Most cities have an OR below 1, indicating that homicides involving male victims are generally less likely to be solved compared to female victims.


# Problem 3

```{r}
birth_weight = read_csv("data/birthweight.csv")

birth_weight = birth_weight |>
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    mrace = as.factor(mrace),
    malform = as.factor(malform)
  )


sum(is.na(birth_weight))

summary(birth_weight) |>
  knitr::kable()
```

- Firstly, do the data cleaning step, including convert numeric variables to factors, checking for missing data etc. 

- In modeling process, started with known factors that are hypothesized to affect birthweight like `delwt`, `gaweeks`, `smoken`, `fincome` and `frace`, these variables were selected based on biological and social theories of birth outcomes.

- Use 
```{r}
birthweight_model = lm(bwt ~ delwt + gaweeks + smoken + fincome + frace, data = birth_weight)

birth_weight = birth_weight |>
  add_predictions(birthweight_model, var = "fitted") |>
  add_residuals(birthweight_model, var = "residuals")

ggplot(birth_weight, aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()
```

```{r}
cv_df = crossv_mc(birth_weight, 100) |> 
  mutate(
    train = map(train, as_tibble), 
    test = map(test, as_tibble)
  )

cv_results = cv_df |>
  mutate(
    my_model = map(train, ~ lm(bwt ~ babysex + bhead + blength + delwt + fincome + 
                                   gaweeks + mheight + mrace + parity + ppwt + smoken, data = .x)),
    model_main = map(train, ~ lm(bwt ~ gaweeks + blength, data = .x)),
    model_interaction = map(train, ~ lm(bwt ~ bhead * blength * babysex, data = .x))
  ) |> 
  mutate(
    rmse_my_model = map2_dbl(my_model, test, ~rmse(model = .x, data = .y)),
    rmse_model_2 = map2_dbl(model_main, test, ~rmse(model = .x, data = .y)),
    rmse_model_3 = map2_dbl(model_interaction, test, ~rmse(model = .x, data = .y))
  )

cv_results_long = cv_results |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |>
  mutate(model = fct_inorder(model))

ggplot(cv_results_long, aes(x = model, y = rmse, fill = model)) +
  geom_boxplot() +
  labs(
    title = "Comparison of Model RMSEs",
    x = "Model",
    y = "Root Mean Squared Error (RMSE)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```
- `my_model` has the lowest RMSE than `main_effect` model and `interaction` model.
